---
title: "Glucocorticoid receptor dynamica"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Glucocorticoids (GC) are a type of steroid hormones which are used by the immune system to repress inflammation. For this reason GCs are used as a medicine to treat inflammation in inflammatory diseases like asthma. The anti-inflammatory process caused by GCs can be divided into two processes: trans-repression and trans-activation. In trans-repression some inflammatory genes are suppressed by binding of the GCs to GC receptors (GR). This causes histon deacetylation of these genes and this causes for less transcription. In trans-activation the GRs activate transcription of anti-inflammatory genes by binding to DNA recognition sites on these genes. This activation is achieved by histon acetylation.

## Receptor-medicine interaction

When the MPL (D) is added it binds to the receptors (R) to form MPL-receptor complexes, leading to an increase in MPL-receptor density (DR). This DR is then being transferred to the nucleus (DR(N)), resulting in a higher DR(N) concentration. However, when the DR(N) concentration reaches a certain level it supresses the mRNA synthesis (the mRNA decreases to 50% of the base value). This then results in less R and therefor decreasing values of DR and DR(N).

## Model
```{r}
library(deSolve)

mRNA.dynamic <- function(t, state, paramters) {
  with(as.list(c(state, paramters)),{
    dmRNAr <- ks.Rm * (1 - DR.N / (IC50.Rm + DR.N)) - k.dRm * mRNAr
    
    dR <- k.sR * mRNAr + Rf * k.re * DR.N - k.on * D * R - k.dR * R
    
    dDR <- k.on * D * R - k.T * DR
    
    dDR.N <- k.T * DR - k.re * DR.N
    
    list(c(dmRNAr, dR, dDR, dDR.N))
  }) 
}

parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

times <- seq(0, 48)

out <- ode(y = state, times = times, func = mRNA.dynamic, parms = parameters)
plot(out, las = 1)
```

## Result discussion

mRNAR is the concentration of the mRNA receptor, at the start we see a high concentration, then it rapidly decrease, after that it increases a bit again. After a certain time it is constant. 
R is the free glucocorticoid receptor density in the cytosol, it has an initial value and increases over time, after reaching a maximum, it decreases. Then after a certain time it is constant.
DR is the density of the MPL-receptor complex and DR.N the concentration of MPL-receptor complex in the cel nucleus. Both start at zero, increase over time and after reaching a maximum it decreases. Then both variables become constant. 

mRNAR is available in high concentration, because of this there is a high concentration of receptors and in a short period of time MPL-receptor complexes are made. This explains the steep increase of DR and DR.N. DR.N supresses the synthesis of mRNAR. Meanwhile the concentration of mRNAR decreases. At some point the mRNAR concentration is less than 50% of the initial value. DR.N doesn't need to inhibit the synthesis of mRNAR anymore, so the concentration of DR.N decreases. Then mRNAR is synthesized until it is 50% of the initial value, after that an equilibrium is established in all 4 variables.

At last, the DR(N) is probably the most import variable for the medicine to work. This is because the medicine needs to get inside the nucleus in order to perform its function. When it is outside of the nucleus (inside the cytosol) it can not reach the genes and interact with it to influence the gene expression. 

