---
title: "Glucocorticoid receptor dynamica"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Glucocorticoids (GC) are a type of steroid hormones which are used by the immune system to repress inflammation. For this reason GCs are used as a medicine to treat inflammation in inflammatory diseases like asthma. The anti-inflammatory process caused by GCs can be divided into two processes: trans-repression and trans-activation. In trans-repression some inflammatory genes are supressed by binding of the GCs to GC receptors (GR). This causes histon deacetylation of these genes and this causes for less transcription. In trans-activation the GRs activate transcription of anti-inflammatory genes by binding to DNA recognition sites on these genes. This activation is achieved by histon acetylation


```{r}
library(deSolve)

mRNA.dynamic <- function(t, state, paramters) {
  with(as.list(c(state, paramters)),{
    dmRNAr <- ks.Rm * (1 - DR.N / (IC50.Rm + DR.N)) - k.dRm * mRNAr
    
    dR <- k.sR * mRNAr + Rf * k.re * DR.N - k.on * D * R - k.dR * R
    
    dDR <- k.on * D * R - k.T * DR
    
    dDR.N <- k.T * DR - k.re * DR.N
    
    list(c(dmRNAr, dR, dDR, dDR.N))
  }) 
}

parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

times <- seq(0, 48)

out <- ode(y = state, times = times, func = mRNA.dynamic, parms = parameters)
plot(out)
```
