---
title: "Glucocorticoid receptor dynamica"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Assignment 1: assess model validity

##### [1] Why is it best practice to plot the median for the experimental data? Explain in your report

The median is less senstive for possible outliers. If a certain value is very high or very low this does not affect the value of the median. When compared to the mean, for example, this does get affected by an outlier.

##### [2] How do the results of the simulations depend on the dose and concentration of the drug? Compare the model variables mRNA, R with the experimental data by running the simulations adjusting dosis D and plot these to find the answer.

```{r}
library(deSolve)

corticosteroid <- function(t, state, paramters) {
  with(as.list(c(state, paramters)),{
    dmRNAr <- ks.Rm * (1 - DR.N / (IC50.Rm + DR.N)) - k.dRm * mRNAr
    
    dR <- k.sR * mRNAr + Rf * k.re * DR.N - k.on * D * R - k.dR * R
    
    dDR <- k.on * D * R - k.T * DR
    
    dDR.N <- k.T * DR - k.re * DR.N
    
    list(c(dmRNAr, dR, dDR, dDR.N))
  }) 
}

parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

times <- seq(0, 48)

out.normal <- ode(y = state, times = times, func = corticosteroid, parms = parameters)

parameters['D'] <- 80 * 1000 * 1/374.471 
out.high <- ode(y = state, times = times, func = corticosteroid, parms = parameters)

parameters['D'] <- 1 * 1000 * 1/374.471 
out.low <- ode(y = state, times = times, func = corticosteroid, parms = parameters)

plot(out.normal[,1], out.normal[,2], las = 1, main = "mRNAr", ylim = c(0, 5), type = "l",
     xlab = "time (h)", ylab = "receptor mRNA")
lines(out.high[,2], col = "red")
lines(out.low[,2], col = "green")
legend(x = 35, y = 2, legend = c("low dosis", "normal dosis", "high dosis"), fill = c("green", "black", "red"))

plot(out.normal[,1], out.normal[,3], las = 1, main = "R", ylim = c(0, 275), type = "l",
     xlab = "time (h)", ylab = "free receptor density ")
lines(out.high[,3], col = "red")
lines(out.low[,3], col = "green")
legend(x = 35, y = 200, legend = c("low dosis", "normal dosis", "high dosis"), fill = c("green", "black", "red"))
```


mRNAr: The concentration of mRNAr is influenced by the MPL concentration. When a lower MPL concentration is used the mRNAr is decreasing slower. The DR(N) affects the mRNAr synthesis, this DR(N) concentration is lower when there is less MPL and therefore the mRNAr does not decrease as much. When a higher dosis is used the mRNAr is decreasing more rapidly and a lower end concentration at equilibrium can be seen. This is becuase the medicine can bind to more receptors, which results in more DR and therefore more DR(N). The IC50 value is reached earlier and the mRNAr concentration gets decreased.


R: The R concentration decreases less when using a lower MPL concentration. This is due to the fact that there is less MPL to create a MPL-receptor complex, so the R concentration stays high. The opposite can be seen when using a higher MPL concentration. The receptor concentration decreases very fast and has a lower end concentration. This is because there is more medicine to bind to the receptor, resulting in a lower receptor concentration.


##### [3] Are the results of the model in line with experimental data? If not, what could be the reason? Think of at least one explanation. Try to test it with simulations (you will get bonus points for that, your explanation does not need to be correct, but should be logical).

```{r}
MPL <- read.csv("MPL.csv")
```

```{r}
dose.0.1 <- MPL[MPL$dose == 0.1,]
dose.0.3 <- MPL[MPL$dose == 0.3,]

medians <- aggregate(MPL[,c("MPL_conc","mRNA","Free_receptor")],list(MPL$dose,MPL$time), median, na.rm=T)
names(medians)[1:2] <- c("dose","time")

median.0.1 <- medians[medians$dose == 0.1,]
median.0.3 <- medians[medians$dose == 0.3,]

# Changes states of model from medians on time = 0
state['mRNAr'] <- medians[medians$time == 0,]$mRNA
state['R'] <- medians[medians$time == 0,]$Free_receptor

# Change the duration of the model with time points from experimental data
times <- unique(MPL$time)

# Change dosis concentration parameter (D) for 0.1 dosis
median_MPL_01 <- median(MPL$MPL_conc[MPL$dose==0.1], na.rm=T)
parameters['D'] <- median_MPL_01 * 1000 * 1/374.471

# Simulate model for 0.1 dosis
out.0.1 <- ode(y = state, times = times, func = corticosteroid, parms = parameters)

# Change dosis concentration parameter (D) for 0.3 dosis
median_MPL_03 <- median(MPL$MPL_conc[MPL$dose==0.3], na.rm=T)
parameters['D'] <- median_MPL_03 * 1000 * 1/374.471

# Simulate model for 0.3 dosis
out.0.3 <- ode(y = state, times = times, func = corticosteroid, parms = parameters)

par(mfrow = c(2,2), cex = 0.8)

# Plot mRNA plot for 0.1 dose
plot(dose.0.1$mRNA ~ dose.0.1$time,
     xlab = "time", ylab = "mRNAr concentration", main = "mRNAr plot of dosis 0.1", las =1)
points(median.0.1$mRNA ~ median.0.1$time, col = "red", type = "l")
lines(out.0.1[,1], out.0.1[,2])

# Plot Free Receptor points for 0.1 dose
plot(dose.0.1$Free_receptor ~ dose.0.1$time,
     xlab = "time", ylab = "Free Receptor concentration", main = "Free Receptor plot of dosis 0.1", las = 1)
points(median.0.1$Free_receptor ~ median.0.1$time, col = "red", type = "l")
lines(out.0.1[,1], out.0.1[,3])

# Plot mRNA plot for 0.3 dose
plot(dose.0.3$mRNA ~ dose.0.3$time,
     xlab = "time", ylab = "mRNAr concentration", main = "mRNAr plot of dosis 0.3", las =1 )
points(median.0.3$mRNA ~ median.0.3$time, col = "red", type = "l")
lines(out.0.3[,1], out.0.3[,2])

# Plot Free Receptor points for 0.3 dose
plot(dose.0.3$Free_receptor ~ dose.0.3$time,
     xlab = "time", ylab = "Free Receptor concentration", main = "Free Receptor plot of dosis 0.3", las = 1)
points(median.0.3$Free_receptor ~ median.0.3$time, col = "red", type = "l")
lines(out.0.3[,1], out.0.3[,3])
```

The results of the model are not in line with experimental data, but this differs between the plots. The free receptor plots (both doses) are somewhat similar of that of the model, except for one outlier at a time of 50. The mRNAr plots show way less similarities, where the experimental line fluctuates a lot. 

This may be caused by the structure of the data. Firstly, the data is divided in a strange way. In the beginning there are a lot datapoints and at a higher time less. Secondly, there are not many datapoints per time unit to reliable fit the data. Lastly, the experimental data may contain some bias due to measurement errors or other factors that can influence the concentration of mRNAr or free receptor.

## Assignment 2: simulate scenario’s

[1] What would be the time course concentration of the activated drug-receptor complex if there was no auto-regulation of glucocorticoid receptor, i.e. if there was not effect of drug on the synthesis of the receptor mRNA? What formula needs to be changed? Adjust the model, run the simulation and plot the results to find out.

```{r}
library(deSolve)

corticosteroid <- function(t, state, paramters) {
  with(as.list(c(state, paramters)),{
    dmRNAr <- ks.Rm * (1 - DR.N / (IC50.Rm + DR.N)) - k.dRm * mRNAr
    
    dR <- k.sR * mRNAr + Rf * k.re * DR.N - k.on * D * R - k.dR * R
    
    dDR <- k.on * D * R - k.T * DR
    
    dDR.N <- k.T * DR - k.re * DR.N
    
    list(c(dmRNAr, dR, dDR, dDR.N))
  }) 
}

parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

times <- seq(0, 48)

out <- ode(y = state, times = times, func = corticosteroid, parms = parameters)
plot(out, las = 1)
```

[2] What is the time course of receptor and mRNA concentrations when the drug treatment is stopped? So After the steady state is reached (at time t_steady), D should be set to zero and the simulation should continue from time t_steady till the new steady state is reached (t_steady_second). Run the simulations and plot the results from t = 0 till t_steady_second.

```{r}
parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

t.steady <- seq(0, 35)
t.steady.second <- seq(35, 100)

out.steady <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['D'] <- 0
out.steady.second <- ode(y = tail(out.steady, 1)[,-1], times = t.steady.second, func = corticosteroid, parms = parameters)


plot(out.steady[,2] ~ out.steady[,1], las = 1, xlim = c(0, 100), type = "l", ylab = "mRNAr concentration", xlab = "time", main = "mRNAr")
lines( out.steady.second[,2] ~ out.steady.second[,1])

plot(out.steady[,3] ~ out.steady[,1], las = 1, xlim = c(0, 100), type = "l", ylab = "R concentration", xlab = "time", main = "R")
lines( out.steady.second[,3] ~ out.steady.second[,1])
```

[3] Different corticosteroids show different association rates from receptors (kon) and different dissociation rates (in this model reflected by kre). Assuming the same concentrations of the drug, what is the effect of different values of kon and kre (consider 2 and 5 times increase and decrease of both parameters separately) on the receptor and mRNA dynamics? Adjust kon and kre as below and plot the results of the simulation for each change. Note: Simulations should be run for 4 new values of kon: 0.00329/5, 0.00329/2, 0.00329*2 and 0.00329*5. The results should be compared to the basic scenario when kon=0.00329 Separately, simulations should be run for 4 new values of kre: 0.57/5, 0.57/2, 0.57*2 and 0.57*5. The results should be compared to the basic scenario when kre= 0.57

```{r}
library('deSolve')

corticosteroid <- function(t, state, paramters) {
  with(as.list(c(state, paramters)),{
    dmRNAr <- ks.Rm * (1 - DR.N / (IC50.Rm + DR.N)) - k.dRm * mRNAr
    
    dR <- k.sR * mRNAr + Rf * k.re * DR.N - k.on * D * R - k.dR * R
    
    dDR <- k.on * D * R - k.T * DR
    
    dDR.N <- k.T * DR - k.re * DR.N
    
    list(c(dmRNAr, dR, dDR, dDR.N))
  }) 
}

parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

t.steady <- seq(0, 48)

kon.normal <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.on'] <- 0.00329 * 2
kon.increase2 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.on'] <- 0.00329 * 5
kon.incerease5 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.on'] <- 0.00329 / 2
kon.decrease2 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.on'] <- 0.00329 / 5
kon.decrease5 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)


parameters['k.on'] <- 0.00329

kre.normal <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.re'] <- 0.57 * 2
kre.increase2 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.re'] <- 0.57 * 5
kre.incerease5 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.re'] <- 0.57 / 2
kre.decrease2 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.re'] <- 0.57 / 5
kre.decrease5 <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
```

```{r}
par(mfrow = c(2,2))
plot(kon.normal[,2] ~ kon.normal[,1], type = 'l', ylim = c(0, 5),
     xlab = "time", ylab = "mRNAr concentration", main = "mRNAr when k.on is changed")
lines(kon.increase2[,2], col = 'red')
lines(kon.incerease5[,2], col = 'green')
lines(kon.decrease2[,2], col = 'blue')
lines(kon.decrease5[,2], col = "orange")

plot(kon.normal[,3] ~ kon.normal[,1], type = 'l',ylim = c(0, 300),
     xlab = "time", ylab = "free receptor", main = "R when k.on is changed")
lines(kon.increase2[,3], col = 'red')
lines(kon.incerease5[,3], col = 'green')
lines(kon.decrease2[,3], col = 'blue')
lines(kon.decrease5[,3], col = "orange")

plot(kre.normal[,2] ~ kon.normal[,1], type = 'l', ylim = c(0, 5),
     xlab = "time", ylab = "mRNAr concentration", main = "mRNAr when k.er is changed")
lines(kre.increase2[,2], col = 'red')
lines(kre.incerease5[,2], col = 'green')
lines(kre.decrease2[,2], col = 'blue')
lines(kre.decrease5[,2], col = "orange")

plot(kre.normal[,3] ~ kon.normal[,1], type = 'l',ylim = c(0, 300),
     xlab = "time", ylab = "free receptor", main = "R when k.er is changed")
lines(kre.increase2[,3], col = 'red')
lines(kre.incerease5[,3], col = 'green')
lines(kre.decrease2[,3], col = 'blue')
lines(kre.decrease5[,3], col = "orange")

legend(x = -22, y = 620, legend = c("normal", "increase2", "incease5", "decrease2", "decrease5"), 
       fill = c("black", "red", "green", "blue", "orange"), xpd = NA, cex = 0.9)
```

k.on is the rate of the formation of the medicine-receptor complex
k.er is the rate of the translocation of the receptor from the nucleus to the cytosol

* mRNAr when k.on is changed
When the k.on rate is increased, the mRNAr concentration decreases because k.on causes an increase in medicine-receptor complexes, which in turn increases more medicine-receptor complexes in the nucleus. And this increases leads to reaching faster of the IC50 value, which lowers the mRNAr synthesis. When the k.on rate is decreased, the mRNAr concentration increases because a lower k.on causes an decrease in medicine-receptor complexes, which in turn decreases the amount of medicine-receptor complexes in the nucleus. And this decrease leads to reaching of the IC50 value slower, which increases the mRNAr synthesis. 

* mRNAr when k.er is changed
When the k.er is increased, the mRNAr concentration increases because the medicine-receptor concentration decreases in the nucleus, which leads to a slower reaching of the IC50 value. If the IC50 is reached slower, the mRNAr synthesis is higher. When the k.er is decreased, the mRNAr concentration decreases because the medicine-receptor concentration increases in the nucleus, which leads to a facter reaching of the IC50 value. If the IC50 is reached faster, the mRNAr synthesis is inhibited. 

* R when k.on is changed
When the k.on rate is increased, the R concentration decreases because more medicine-receptor complexes are being formed. When the k.on is decrease, the R concentration increases because less medicine-receptor are being fomred.

* R when k.er is changed
When the k.er rate is increases, the R concentration increases because more receptors are being translocated from the nucleus to the cytosol. When the k.er rate is decreased, the R concentration decreases because less receptort are being translocated from the nucleus to the cytosol.


[4] What would happen if the synthesis of the receptor was completely blocked? Which parameter needs to be put to zero? Adjust the parameter, run the simulations and plot the results.

The ks_r needs to be put to zero, because this is the rate of the receptor synthesis
```{r}
parameters <- c(ks.Rm = 2.90,
                IC50.Rm = 26.2,
                k.on = 0.00329,
                k.T = 0.63,
                k.re = 0.57,
                Rf = 0.49,
                k.dR = 0.0572,
                k.dRm = 0.612,
                k.sR = 3.22,
                D = 20 * 1000 * 1/374.471 )

state <- c(mRNAr = 4.74,
           R = 267,
           DR = 0,
           DR.N = 0)

t.steady <- seq(0, 48)

out.normal <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)
parameters['k.sR'] <- 0
out.ksr <- ode(y = state, times = t.steady, func = corticosteroid, parms = parameters)


plot(out.normal[,2] ~ out.normal[,1], las = 1, type = "l", ylab = "mRNAr concentration", xlab = "time", main = "mRNAr")
lines(out.ksr[,2] ~ out.ksr[,1], col= "red")

plot(out.normal[,3] ~ out.normal[,1], las = 1, type = "l", ylab = "R concentration", xlab = "time", main = "R")
lines(out.ksr[,3] ~ out.ksr[,1], col = "red")

plot(out.normal[,4] ~ out.normal[,1], las = 1, type = "l", ylab = "DR concentration", xlab = "time", main = "DR")
lines(out.ksr[,4] ~ out.ksr[,1], col = "red")

plot(out.normal[,5] ~ out.normal[,1], las = 1,type = "l", ylab = "DR(N) concentration", xlab = "time", main = "DR(N)")
lines(out.ksr[,5] ~ out.ksr[,1], col = "red")
```





